---
title: "Hands-on_Ex10"
author: "Luo Yuming"
date: "Nov 01 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
---

------------------------------------------------------------------------

# Processing and Visualising Flow Data

# 15.1 Overview

Spatial interaction represent the flow of people, material, or information between locations in geographical space. It encompasses everything from freight shipments, energy flows, and the global trade in rare antiquities, to flight schedules, rush hour woes, and pedestrian foot traffic.

Each spatial interaction, as an analogy for a set of movements, is composed of a discrete origin/destination pair. Each pair can be represented as a cell in a matrix where rows are related to the locations (centroids) of origin, while columns are related to locations (centroids) of destination. Such a matrix is commonly known as an origin/destination matrix, or a spatial interaction matrix.

In this hands-on exercise, you will learn how to build an OD matrix by using Passenger Volume by Origin Destination Bus Stops data set downloaded from LTA DataMall. By the end of this hands-on exercise, you will be able:

to import and extract OD data for a selected time interval, to import and save geospatial data (i.e. bus stops and mpsz) into sf tibble data frame objects, to populate planning subzone code into bus stops sf tibble data frame, to construct desire lines geospatial data from the OD data, and to visualise passenger volume by origin and destination bus stops by using the desire lines data.

# 15.2 Getting Started

```{r}
pacman::p_load(tmap, sf, DT, stplanr, tidyverse)
```

# 15.3 Preparing the Flow Data

## 15.3.1 Importing the OD data

```{r}
odbus <- read_csv("data/aspatial/origin_destination_bus_202210.csv")
```

```{r}
glimpse(odbus)
```

```{r}
odbus$ORIGIN_PT_CODE <- as.factor(odbus$ORIGIN_PT_CODE)
odbus$DESTINATION_PT_CODE <- as.factor(odbus$DESTINATION_PT_CODE) 
```

## 15.3.2 Extracting the study data

For the purpose of this exercise, we will extract commuting flows on weekday and between 6 and 9 o’clock.

```{r}
odbus6_9 <- odbus %>%
  filter(DAY_TYPE == "WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 6 &
           TIME_PER_HOUR <= 9) %>%
  group_by(ORIGIN_PT_CODE,
           DESTINATION_PT_CODE) %>%
  summarise(TRIPS = sum(TOTAL_TRIPS))
```

```{r}
datatable(odbus6_9)
```

We will save the output in rds format for future used.

```{r}
write_rds(odbus6_9, "data/rds/odbus6_9.rds")
```

```{r}
odbus6_9 <- read_rds("data/rds/odbus6_9.rds")
```

# 15.4 Working with Geospatial Data

For the purpose of this exercise, two geospatial data will be used. They are:

BusStop: This data provides the location of bus stop as at last quarter of 2022. MPSZ-2019: This data provides the sub-zone boundary of URA Master Plan 2019. Both data sets are in ESRI shapefile format.

## 15.4.1 Importing geospatial data

```{r}
busstop <- st_read(dsn = "data/geospatial",
                   layer = "BusStop") %>%
  st_transform(crs = 3414)

mpsz <- st_read(dsn = "data/geospatial",
                   layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```

```{r}
mpsz
mpsz <- write_rds(mpsz, "data/rds/mpsz.rds")
```

# 15.5 Geospatial data wrangling

## 15.5.1 Combining Busstop and mpsz

```{r}
busstop_mpsz <- st_intersection(busstop, mpsz) %>%
  select(BUS_STOP_N, SUBZONE_C) %>%
  st_drop_geometry()
datatable(busstop_mpsz)
```

```{r}
write_rds(busstop_mpsz, "data/rds/busstop_mpsz.rds")  
```

Next, we are going to append the planning subzone code from busstop_mpsz data frame onto odbus6_9 data frame.

```{r}
od_data <- left_join(odbus6_9 , busstop_mpsz,
            by = c("ORIGIN_PT_CODE" = "BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_SZ = SUBZONE_C,
         DESTIN_BS = DESTINATION_PT_CODE)
```

Before continue, it is a good practice for us to check for duplicating records.

```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```

If duplicated records are found, the code chunk below will be used to retain the unique records.

```{r}
od_data <- unique(od_data)
```

```{r}
od_data <- left_join(od_data , busstop_mpsz,
            by = c("DESTIN_BS" = "BUS_STOP_N")) 
```

```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```

```{r}
od_data <- od_data %>%
  rename(DESTIN_SZ = SUBZONE_C) %>%
  drop_na() %>%
  group_by(ORIGIN_SZ, DESTIN_SZ) %>%
  summarise(MORNING_PEAK = sum(TRIPS))
```

```{r}
write_rds(od_data, "data/rds/od_data_fii.rds")
```

```{r}
od_data_fii <- read_rds("data/rds/od_data.rds")
```

# 15.6 Visualising Spatial Interaction

## 15.6.1 Removing intra-zonal flows

```{r}
od_data_fij <- od_data[od_data$ORIGIN_SZ!=od_data$DESTIN_SZ,]
write_rds(od_data_fij, "data/rds/od_data_fij.rds")
od_data_fij <- read_rds("data/rds/od_data_fij.rds")
```

## 15.6.2 Creating desire lines

```{r}
flowLine <- od2line(flow = od_data_fij, 
                    zones = mpsz,
                    zone_code = "SUBZONE_C")
write_rds(flowLine, "data/rds/flowLine.rds")
flowLine <- read_rds("data/rds/flowLine.rds")
```

## 15.6.3 Visualising the desire lines

```{r}
tm_shape(mpsz) +
  tm_polygons() +
flowLine %>%  
tm_shape() +
  tm_lines(lwd = "MORNING_PEAK",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           alpha = 0.3)
```

When the flow data are very messy and highly skewed like the one shown above, it is wiser to focus on selected flows, for example flow greater than or equal to 5000 as shown below.

```{r}
tm_shape(mpsz) +
  tm_polygons() +
flowLine %>%  
  filter(MORNING_PEAK >= 5000) %>%
tm_shape() +
  tm_lines(lwd = "MORNING_PEAK",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           alpha = 0.3)
```
